---
title: Supplementary references for “Colonial history and global economics distort
  our understanding of deep-time biodiversity”
corresauth: '*Corresponding author, email: nussaibah.raja.schoob@fau.de'
abstract: |
  This document contains the references for the data sources used in the study:
  
  Raja, N. B., Dunne, E. M., Matiwane, A., Khan, T. M., Nätscher, P. S., Ghilardi, A. M., and Chattopadhyay, D. (2021). Colonial history and global economics distort our understanding of deep-time biodiversity. Nature Ecology & Evolution. https://doi.org/10.1038/s41559-021-01608-8
  
keywords: paleontology, biodiversity, sampling biases, scientific colonialism, research ethics, scientometrics
author:
- name: Nussaïbah B. Raja*
  affiliation: GeoZentrum Nordbayern, Department of Geography and Geosciences, Friedrich-Alexander
    University Erlangen-Nürnberg, Erlangen, Germany
- name: ' Emma Dunne'
  affiliation: '  School of Geography, Earth and Environmental Sciences, University
    of Birmingham, Edgbaston, Birmingham, B15 2TT, UK'
- name: ' Aviwe Matiwane'
  affiliation: ' Department of Botany, Rhodes University, 3230 Lucas Ave, Grahamstown,
    6139, South Africa'
- name: |2-
     Tasnuva Ming
    Khan
  affiliation: ' GeoZentrum Nordbayern, Department of Geography and Geosciences, Friedrich-Alexander
    University Erlangen-Nürnberg, Erlangen, Germany'
- name: ' Paulina S. Nätscher'
  affiliation: ' GeoZentrum Nordbayern, Department of Geography and Geosciences, Friedrich-Alexander
    University Erlangen-Nürnberg, Erlangen, Germany'
- name: ' Aline M. Ghilardi'
  affiliation: ' Federal University of Rio Grande do Norte (UFRN), Department of Geology,
    Natal, RN, Brazil'
- name: ' Devapriya Chattopadhyay'
  affiliation: '  IISER Pune, Department of Earth and Climate Science, Dr. Homi Bhaba
    Road, Pune 411008, India'
params:
  source: The Paleobiology Database
  download url for occurrences: https://paleobiodb.org/data1.2/occs/list.csv?datainfo&rowcount&interval=Ediacaran,Holocene&show=class,classext,genus,subgenus,coll,coords,loc,paleoloc,strat,stratext,lith,env,ref,crmod,timebins,timecompare
  download url for references: https://paleobiodb.org/data1.2/occs/refs.csv?interval=Ediacaran,Holocene&select=occs&show=formatted,entname
  date_downloaded: 19/01/2021
  data_format: Processed
  data_access: Available to download from https://dx.doi.org/10.17605/OSF.IO/6WC7A
  location: Global
  time: 1990 - 2020
enterer_names: enterernames.txt
bibliography:
- references.bib
- packageref.bib
- missing.bib

output:
  pdf_document:
    citation_package: biblatex
    keep_tex: no
    fig_caption: yes
    latex_engine: xelatex
    template: template.tex
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 10pt
#feel free to change to accomodate for other styles, e.g. alphabetic
biblio-style: numeric
---
# Overview

```{r echo=FALSE, message=FALSE, warning=FALSE}
#create data frame and formats text based on parameters provided
specs <- gsub("_", " ", refer::capitalize(names(params)))
details <- as.character(unlist(params))

df <- data.frame(Specifications=trimws(specs),
                Details=refer::wrap_url(details, 52))


kk <- knitr::kable(df, "latex", booktabs=TRUE, col.names = NULL,
                   caption="Specifications Table")

# #makes first column bold
kk <- kableExtra::column_spec(kk, 1, bold=TRUE)
# 
#increases the width of second column and allows wrapping of text
kk <- kableExtra::column_spec(kk, 2, width = "25em")
# 
# tell latex to keep the position of the table
kk <- kableExtra::kable_styling(kk, latex_options="hold_position")

kk
```

# Data Processing

The affiliate countries of the authors of each of the publications were compiled using (1) webscraping methods and (2) manually when the former was not possible. When the publications could not be found or the metadata could not be found, alternative sources such as personal websites, academic social networks such as ResearchGate or Academia.edu, or published obituaries (in the case when one or more of the authors were deceased). The affiliate dataset was the merged with the occurrences dataset of the Paleobiology Database to link the affiliate countries to where sampling took place. All countries were standardised using the official three-letter country codes (`ISO 3166-1 alpha-3`) the International Organization for Standardization (ISO, https://www.iso.org/) using the R package `countrycode` `r packageVersion("countrycode")`[@countrycode2018].

# Data description

Sample dataset:
```{r sample_data, echo=FALSE}
read.csv("sample_data.csv")

```
The final dataset comprises 5 columns:

*     `reference_no`: the id of the publication as per the Paleobiology Database
*     `samp_country`: the country where sampling took place
*     `aff_country`: the affiliate country of the author(s) of the publication
*     `samp_code`: the country code for the sampling country
*     `aff_code`: the country code for the affiliate country

Each row in the reference represents one affiliate country and sampling country. As such, in some cases, one reference can have multiple rows, e.g. 74895 where the authors of the publication were from Russia but the data was obtained from the UK, Japan and Russia. The first occurring row of a reference contains the affiliate country of the first author.

# Acknowledgements
We would like to thank John Alroy and Charles Marshall, in addition to A. Miller, M. Koznik and others for their vision and critical early work in establishing the Paleobiology Database as well as the [numerous contributors of data](#enterers), without all of whom this study would not have been possible. This report was generated using the R package `refer` `r packageVersion("refer")` [@refer2021].
